Global land characterisation using cover fractions at 100m resolution

### 2020-07-22 article/article.tex
Resolving most comments of draft 2

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS-LC100}{CGLS-LC100}{Copernicus Global Land Services 100 m Land Cover}
\newacronym{CGLS}{CGLS}{Copernicus Global Land Services}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{IIASA}{IIASA}{International Institute for Applied Systems Analysis}
\newacronym{WUR}{WUR}{Wageningen University \& Research}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym{OSAVI}{OSAVI}{Optimised Soil-Adjusted Vegetation Index}
\newacronym{NDVI}{NDVI}{Normalised Difference Vegetation Index}
\newacronym{NDMI}{NDMI}{Normalised Difference Moisture Index}
\newacronym{NIRv}{NIRv}{\gls{NIR} of vegetation}
\newacronym{NIR}{NIR}{near infra-red}
\newacronym{SCM}{SCM}{subpixel confusion-uncertainty matrix}
\newacronym{FNC}{FNC}{fuzzy nearest centroid}
\newacronym{PLS}{PLS}{partial least squares}
\newacronym{MLR}{MLR}{multinomial logistic regression}
\newacronym{RF}{RF}{random forest}
\newacronym{TOC}{TOC}{Top-of-Canopy}
\newacronym{CART}{CART}{Classification and regression tree}
\newacronym{GLM}{GLM}{General linear regression model}
\newacronym{LCCS}{LCCS}{UN Land Cover Classification Scheme}
%\makenoidxglossaries
% Disable abbreviating RF for now [text={random forest}]
%\glsunset{RF}

@@

We compared the performance of three classes of machine learning regression algorithms for land cover fraction mapping, for the best model assessed the importance of six types of training variables, evaluated the spatial and per-class accuracy, and demonstrated a wall-to-wall prediction of seven land cover fractions over the globe at a sampling interval of 0.2 degrees.

The models were trained on over 130\,000 points and validated on a separate dataset of over 20\,000 points of the Copernicus Global Land System: Dynamic Land Cover (CGLS-LC100) project. Both datasets are global and aligned with the Proba-V 100 m grid.

One challenge for fraction mapping models is data sparsity: most fractions are 0\%, but regression favours the mean, thus 0\% and 100\% are difficult to predict well. We proposed a new solution by combining three models. A binary model determines whether a pixel is pure; if so, it is processed using a classification model; otherwise with a regression model.

Results showed that from the compared models, random forest regression reached the lowest RMSE of 17.3\%. Lowest MAE (7.9\%) and highest overall accuracy (72±2\%) was achieved by a tuned random forest model with our proposed three-model approach, which results in better estimation of extreme values at the cost of more misclassification. The models were driven primarily by optical remote sensing data, followed by soil and climate data. Herbaceous cover was the most challenging to discern for all models, whereas built-up and water were the easiest to estimate.

This research proves that machine learning algorithms can be applied globally to a wide variety of land cover classes to create an exhaustive global land cover fraction map. This expresses land cover more precisely, and empowers users to create their own discrete maps using user-defined thresholds and rules, which helps to customise the result for a diverse range of uses. In addition, this work contributes to efforts to produce exhaustive global land cover fractions as an operational product, and to assess the accuracy of land cover fraction maps both thematically and spatially.
\end{abstract}

@@

\begin{itemize}
    \item Three groups of statistical and machine learning models tested for deriving seven land cover fractions at a global scale
    \item Random forest regression performed the best (17.3 RMSE or 7.9 MAE)
    \item Remote sensing data was the most important for model predictions with varying features relevant for different fraction predictions
    \item Independent validation results show highest accuracies for built-up and water fractions, lower accuracies are found for grasses and shrubs
    \item Comprehensive multi-fraction global land cover characterisation shown as an alternative to discrete classification
\end{itemize}

@@

Land cover, as one of the key variables for monitoring a number of \glspl{SDG}, has received increasing amounts of attention, also as new satellites are launched and the capacity for land monitoring increases.
In this context, new global land cover maps have emerged to better map the current land cover of the world, as well as to track land cover change.
Some of the recent achievements have been the \ac{LC-CCI} map \citep{defourny2012cci} that aims at producing a long-term set of consistent global annual medium-resolution land cover maps aimed at the climate community, \ac{CGLS-LC100} map \citep{buchhorn_moderate_2019} that aims at a finer spatial resolution and high quality with yearly updates since 2015, and the \ac{FROM-GLC} map \citep{fromglc2019} that showcases the potential of land cover mapping at 10 m resolution.

All of these global land cover maps (with the exception of \ac{CGLS-LC100} fraction layers), as well as previous global land cover maps such as by \citet{bartholome2005glc2000, friedl2010modis, arino2007globcover, see2015hybrid, chen2015globeland30}, are provided as discrete classes (also known as ``hard'' or ``crisp'' classification), where each pixel of the map can only represent a single class.
However, this results in less precise estimation of land cover due to mixed pixels, where in reality there are multiple land cover classes in the area covered by a single pixel.
This issue becomes worse at coarse resolutions and heterogeneous landscapes.
It may result in biases, for instance, a sparse forest may be classified as grassland, ignoring the relatively few trees in the area, and thus ultimately lead to an underestimation of tree cover at large scales.

@@

Land cover fraction mapping has been attempted in the past, both in research \citep{adams_classification_1995, foody1996fuzzyevaluation, colditz_land_2011, sharma_assessing_2011, uma_shankar_wavelet-fuzzy_2011, dwivedi_optimisation_2012, lizarazo_quantitative_2012, gessner_estimating_2013, okujeni_generalizing_2018} and production (e.g. \citealp{hansen2000hardtree, hansen_continuous_2011, pengra_global_2015, hansen_global_2003, sexton_global_2013}).
Nevertheless, to date only \ac{CGLS-LC100} provides a global map with fractions of every major land cover class.

Land cover fraction mapping can be performed using a variety of different approaches and algorithms.
In its core, it is a regression rather than a classification problem, as the output is a fraction of a label rather than a label itself.
Popular approaches include fuzzy nearest centroid regression \citep{zhang2001fullyfuzzy}, \ac{SMA} (e.g. \citealp{yang_landsat_2012, hobbs2003linear, shimabukuro1991least}), \ac{RF} regression \citep{schwieder_estimating_2014}, \gls{SVM} regression \citep{schwieder_estimating_2014}, \gls{MLP} \glspl{NN} \citep{zhang2001fullyfuzzy}, genetic algorithms \citep{stavrakoudis_boosted_2011} and wavelet transformation \citep{uma_shankar_wavelet-fuzzy_2011}.
So far other studies have only compared a few of these methods at once, and never with an exhaustive set of land cover classes nor at global scale.

In addition to classification methods, one aspect that is important for classification accuracy is the data (covariates) that the algorithm is trained on and predict land cover from \citep{yu2014metadiscoveries}. In addition to spectral covariates (blue, red, near infrared, shortwave infrared bands), temporal covariates (growing season start, duration, growing intensity) are commonly used to improve the separability of crops \citep{jakubauskas2001harmonic}. Elevation covariates (elevation, slope, aspect) are known to improve the separability between tree classes \citep{burrough2001fuzzy}. \Glspl{VI} have been used to improve the detection of wetlands \citep{sader1995wetlands}, as well as for defining classification rules in rule-based fuzzy classification \citep{baraldi2006rulebased}. Therefore in this paper such covariates have also been used in order to increase land cover classification accuracy, and their importance for prediction was analysed.

Lastly, accuracy assessment by itself is often times challenging, especially at a global scale.
It requires a comprehensible dataset across the globe that would be comparable with the modelled land cover, and yet independent of the training data used to create it.
For land cover fraction mapping it is even more challenging, as fraction information needs to be available, or alternatively fine spatial resolution data from which the fractions can be calculated.
In this study, we made use of an independent validation set of over 20\,000 points collected across the globe using the same legend, grid and spatial resolution.
The validation data was collected at \gls{WUR}, whereas the training data was collected at \gls{IIASA}, ensuring independence of the datasets.
We used the validation data to assess model accuracy both thematically and spatially.

@@

\begin{enumerate}
 \item Compare the performance of a variety of machine learning regression models for global-scale land cover fraction mapping
 \item Investigate approaches for tuning the models by reducing bias in the predictions with regards to zero inflation and predictions tending towards the mean
 \item Compare the importance of covariates and covariate groups for the prediction of land cover fractions
 \item Assess the accuracy of the models from a thematic and spatial point of view, taking various user needs into account
\end{enumerate}

@@

The reference data (for model training and validation) used in this study was collected as part of the \ac{CGLS} project.
The training set was generated by experts in \ac{IIASA}.
This global point dataset provides information on nine class fractions of sampled pixels of the Proba-V 100 m product.
The fraction estimate was acquired by high-resolution image interpretation by experts, for the year 2015.
Each of the sampled Proba-V pixels was subdivided into a 10-by-10 subpixel grid and the area estimates were derived from the subpixel ratios.

For model validation, an independent global dataset of \ac{CGLS}, generated by \gls{WUR} \citep{tsendbazar_developing_2018}, was used.
The method of point collection is equivalent to the method used for training data collection, but performed independently by a different group of regional experts.

The data includes over 150\,000 training points and over 21\,000 validation points across the globe, describing the fractions of 12 classes in the year 2015.
The classification scheme follows the \gls{LCCS} \citep{buchhorn_moderate_2019}.
However, some of the classes were very detailed and had too few observations to classify well, thus we merged the classes into a total of seven: bare land (including snow and ice), cropland (including shifting cultivation), herbaceous (including wetland, lichen and moss), shrubs, trees, built-up and inland water.
The unknown class was discarded: points with the dominant land cover class marked as unknown were not used.
For points with a minority fraction of unknown, the remaining classes were linearly rescaled to add up to 100\%.
Thus in the end the training dataset size became 138\,164 and the validation dataset size became 20\,705. See figure \ref{fig-reference-data} for the spatial distribution of the points.

@@

\begin{figure}
 \includegraphics[width=\textwidth]{article/article-figures/maps/2020-07-06-training-and-validation}
 \caption{Points representing 100x100 m areas at which land cover reference data was used in the study. The colours represent the dominant land cover class at each point. (a): training dataset, collected by \gls{IIASA}, 138\,164 points used in this study. (b): validation dataset, collected by \gls{WUR}, 20\,705 points used in this study.}
 \label{fig-reference-data}
\end{figure}

@@

% \begin{table}
% \centering
%     \begin{tabular}{llp{6cm}}
%          \toprule
%          \textbf{Category} & \textbf{Number} & \textbf{Data source} \\
%          \midrule
%          Spectral & 22 & Proba-V 100m \ac{TOC} reflectance v1.02 \citep{probavguide2} \\
%          Harmonic & 9 & Proba-V 100m \ac{TOC} reflectance v1.02 \citep{probavguide2} \\
%          Terrain & 4 & ASTER GDEM V003 \citep{ASTGTM003} \\
%          Climate & 21 & WorldClim 2.0 \citep{worldclim2} \\
%          Soil & 8 & SoilGrids \citep{hengl_soilgrids250m_2017} \\
%          Location & 3 & None (intrinsic) \\
%          \bottomrule
%     \end{tabular}
%     \label{tab-inputdata}
%     \caption{Data sources for the covariates input into the model.}
% \end{table}

@@

The entire archive (2014-03-11 to 2019-07-16) of the Proba-V 100 m Level 3 Top-of-Canopy 5-day composite product \citep{probavguide2} was used for this study.
While the reference data corresponded to the land cover at the year 2015 specifically, the whole time series of Proba-V was chosen to obtain more robust statistics for the temporal metrics.
Land cover change is a relatively rare phenomenon and thus is not likely to affect more than 2\% of the training points within the time period.
Nevertheless, this may result in poorer model performance compared to if reference data for each year would be available.

The time series of each of the Proba-V bands were cloud masked, first by applying the status mask provided with the product itself, and then by running a temporal cloud filter to remove the remaining outliers that were further than 2 standard deviations away from the fitted \ac{LOESS} curve over the blue band.
The resulting cloud-free time series was then used to extract descriptive statistics (median and interquartile range) of NDVI, NDMI, EVI, OSAVI and NIR\textsubscript{v} over the whole time series, as well as for each season separately.
Next, harmonic analysis was run on NDVI in order to decompose the time series into sine and cosine components for two frequency orders (annual and semiannual), as well as the trend and intercept of the model.
From this, phase and amplitude for the two orders were calculated.
Lastly, the minimum and maximum of the NDVI time series was also retained as covariates.

% \begin{figure}
%  \includegraphics[width=\textwidth]{article-figures/flowcharts/preprocessing-optical}
%  \caption{Optical data preprocessing chain.}
%  \label{fig-preprocessing-optical}
% \end{figure}

@@

For elevation data, the ASTER GDEM v003 \citep{ASTGTM003} product (30 m) was resampled to Proba-V 100 m grid and used as the elevation covariate.
In addition, GDAL was used to calculate terrain parameters out of elevation: slope, aspect and \ac{TPI}.
%All terrain covariates were then resampled to the Proba-V grid.

@@

For climate data, WorldClim 2.0 30 s product \citep{worldclim2} was used.
It includes monthly temperature, precipitation, solar radiation, wind speed and water vapour pressure data.
In additional, 19 bioclimatic parameters were calculated from the data, using the dismo package \citep{dismo}.
Some additional biophysical parameters were also calculated, namely all of the climate data during the coldest, warmest, driest and wettest months of the year at each location, as well as yearly averages of each of the original parameters.

@@

SoilGrids \citep{hengl_soilgrids250m_2017} was used for getting data about soils.
It is based on a random forest model that predicts soil properties at various soil depths globally at a 250 m resolution.
In the creation of SoilGrids, a land cover map (based on MODIS) had been used, and so, in order to avoid circular inference, the covariates that are significantly influenced by the land cover map as detailed in \citet{hengl_soilgrids250m_2017} were excluded.
The soil taxonomy covariates were also excluded, since they are categorical derivatives from the numerical soil property data and thus do not contribute to land cover fraction prediction.

@@

In total, 313 covariates were used for predicting the land cover fractions.
However, many of them are colinear with one another, and thus variable selection was used to remove colinear covariates.
Covariates that had a Pearson's correlation coefficient $r$ above 0.9 were excluded in an iterative process.
After that, covariates with a Spearman's $\rho$ above 0.9 were likewise excluded.

Most of the colinear covariates were soil covariates at different depths.
Therefore the 10 cm depth covariates were left in, and the other depths excluded, as long as $r$ was above 0.9.
Similarly, climate data was colinear between subsequent months. Thus January and July data was preferred, as these months represent different extremes of the year and are less correlated with other months' data.
In addition, all of the spectral bands of Proba-V were also highly correlated with each other as well as to vegetation indices, so only vegetation indices were kept.

@@

For deriving land cover fractions, we compared a wide array of machine learning regression methods.
They can be broadly divided into three categories.
Tested linear models were \ac{FNC}, \ac{GLM}, \ac{PLS} regression, Lasso regression and \ac{MLR}.
Non-tree machine learning models tested were \ac{MLP} \glspl{NN} and \glspl{SVM}.
Lastly, machine learning models based on \glspl{CART} tested were Cubist and \ac{RF} regressions.
In addition, as a baseline we also compared the results with an intercept model (all fractions always predicted to be equal, namely $\frac{100\%}{7}\approx14.29\%$).

@@

@@

\subsection{Validation/Accuracy Assessment}

@@

\end{document}
\ead{marcel.buchhorn@vito.be}

>>>>

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS-LC100}{CGLS-LC100}{Copernicus Global Land Services 100 m Land Cover}
%\newacronym{CGLS}{CGLS}{Copernicus Global Land Services}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{IIASA}{IIASA}{International Institute for Applied Systems Analysis}
\newacronym{WUR}{WUR}{Wageningen University \& Research}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym{OSAVI}{OSAVI}{Optimised Soil-Adjusted Vegetation Index}
\newacronym{NDVI}{NDVI}{Normalised Difference Vegetation Index}
\newacronym{NDMI}{NDMI}{Normalised Difference Moisture Index}
\newacronym{NIRv}{NIR\textsubscript{v}}{\gls{NIR} of vegetation}
\newacronym{EVI}{EVI}{Enhanced Vegetation Index}
\newacronym{NIR}{NIR}{near infra-red}
\newacronym{SCM}{SCM}{subpixel confusion-uncertainty matrix}
\newacronym{FNC}{FNC}{fuzzy nearest centroid}
\newacronym{PLS}{PLS}{partial least squares}
\newacronym{MLR}{MLR}{multinomial logistic regression}
\newacronym{RF}{RF}{random forest}
\newacronym{TOC}{TOC}{Top-of-Canopy}
\newacronym{CART}{CART}{classification and regression tree}
\newacronym{GLM}{GLM}{General linear regression model}
\newacronym{LCCS}{LCCS}{UN Land Cover Classification Scheme}
\newacronym{GDAL}{GDAL}{Geospatial Data Abstraction Library}
%\makenoidxglossaries
% Disable abbreviating RF for now [text={random forest}]
%\glsunset{RF}

@@

We compared the performance of three types of machine learning regression algorithms for land cover fraction mapping.
For the best model, we assessed the importance of six types of training variables, evaluated the spatial and per-class accuracy, and demonstrated a wall-to-wall prediction of seven land cover fractions over the globe at a sampling interval of 0.2 degrees.
The models were trained on over 138\,000 points and validated on a separate dataset of over 20\,000 points of the \gls{CGLS-LC100} project. Both datasets are global and aligned with the PROBA-V 100 m grid.

One challenge for land cover fraction mapping models is data sparsity: the dataset is zero-inflated, but regression favours the mean, thus 0\% and 100\% are difficult to predict well. We proposed a new solution by combining three models. A binary model determines whether a pixel is pure; if so, it is processed using a classification model; otherwise with a regression model.

Results showed that from the compared models, random forest regression reached the lowest RMSE of 17.3\%. Lowest MAE (7.9\%) and highest overall accuracy (72±2\%) was achieved by a tuned random forest model with our proposed three-model approach. The models were driven primarily by optical remote sensing data, followed by soil and climate data. Herbaceous cover was the most challenging to discern for all models, whereas built-up and water were the easiest to estimate.

This research proves that machine learning algorithms can be applied globally to a wide variety of land cover classes to create an exhaustive global land cover fraction map. Fraction mapping expresses land cover more precisely, and empowers users to create their own discrete maps using user-defined thresholds and rules, which helps to customise the result for a diverse range of uses. In addition, this work directly contributes to the operational production of global land cover fraction layers within the \gls{CGLS-LC100} project, as well as to the accuracy assessment of land cover fraction maps both thematically and spatially.
\end{abstract}

@@

\begin{itemize}
    \item 3 machine learning model groups tested for deriving 7 land cover fractions globally
    \item Random forest regression performed the best (17.3 RMSE or 7.9 MAE)
    \item Optical remote sensing data was the most important input for model predictions
    \item Independent validation showed highest accuracies for built-up and water fractions
    \item Multi-fraction land cover mapping shown as an alternative to discrete classification
\end{itemize}

@@

Land cover, as one of the key variables for monitoring a number of \glspl{SDG}, has received increasing amounts of attention, also as new satellites are launched and the capacity for land monitoring increases.
In this context, new global land cover maps have emerged to better map the current land cover of the world, as well as to track land cover change.
Some of the recent achievements have been the \ac{LC-CCI} map \citep{defourny2012cci} that aims at producing a long-term set of consistent global annual medium-resolution land cover maps aimed at the climate community, \glsentryfull{CGLS-LC100} map \citep{buchhorn_moderate_2019, buchhorn_copernicus_2020} that aims at a finer spatial resolution and high quality with yearly updates since 2015, and the \ac{FROM-GLC} map \citep{fromglc2019} that showcases the potential of land cover mapping at 10 m resolution.

Except for the \ac{CGLS-LC100} product with its cover fraction layers, all other global land cover maps, including previous maps such as by \citet{bartholome2005glc2000, friedl2010modis, arino2007globcover, see2015hybrid, chen2015globeland30}, are provided with discrete classes (also known as ``hard'' or ``crisp'' classification), where each pixel of the map can only represent a single class.
However, this results in less precise estimation of land cover due to mixed pixels, where in reality there are multiple land cover classes in the area covered by a single pixel.
This issue becomes worse at coarse resolutions and heterogeneous landscapes.
It may result in biases, for instance, a sparse forest may be classified as grassland, ignoring the relatively few trees in the area, and thus ultimately lead to an underestimation of tree cover at large scales.

@@

Land cover fraction mapping has been attempted in the past, both in research \citep{adams_classification_1995, foody1996fuzzyevaluation, colditz_land_2011, sharma_assessing_2011, uma_shankar_wavelet-fuzzy_2011, dwivedi_optimisation_2012, lizarazo_quantitative_2012, gessner_estimating_2013, okujeni_generalizing_2018} and production (e.g. \citealp{hansen2000hardtree, hansen_continuous_2011, pengra_global_2015, hansen_global_2003, sexton_global_2013}).
Nevertheless, to date only \ac{CGLS-LC100} \citep{marcel_buchhorn_copernicus_2019} provides global maps with fractions of every major land cover class \citep{nandin-erdene_tsendbazar_copernicus_2019}.

Land cover fraction mapping can be performed using a variety of different approaches and algorithms.
In its core, it is a regression rather than a classification problem, as the output is a fraction of a label rather than a label itself.
Popular approaches include fuzzy nearest centroid regression \citep{zhang2001fullyfuzzy}, \ac{SMA} (e.g. \citealp{shimabukuro1991least, adams_classification_1995, hobbs2003linear, yang_landsat_2012}), \ac{RF} regression \citep{walton2008subpixelrf}, \gls{SVM} regression \citep{walton2008subpixelrf}, Cubist regression \citep{walton2008subpixelrf}, \gls{MLP} \glspl{NN} \citep{zhang2001fullyfuzzy}, genetic algorithms \citep{stavrakoudis_boosted_2011} and wavelet transformation \citep{uma_shankar_wavelet-fuzzy_2011}.
So far other studies have only compared a few of these methods at once, and never with an exhaustive set of land cover classes nor at global scale.

In addition to classification methods, one aspect that is important for classification accuracy is the data (covariates) that the algorithm is trained on and predict land cover from \citep{yu2014metadiscoveries}. In addition to optical covariates (blue, red, near infrared, shortwave infrared spectral bands), temporal covariates (growing season start, duration, growing intensity) are commonly used to improve the separability of crops \citep{jakubauskas2001harmonic}. Elevation covariates (elevation, slope, aspect) are known to improve the separability between tree classes \citep{burrough2001fuzzy}. \Glspl{VI} condense data and reduce the noise inherent in single bands, and thus are often used to improve land cover class separability, e.g. for wetlands \citep{sader1995wetlands}, as well as for defining classification rules in rule-based fuzzy classification \citep{baraldi2006rulebased}. Therefore in this paper such covariates have also been used in order to increase the accuracy of land cover fraction regression model predictions, and their importance for the predictions was analysed.

Lastly, accuracy assessment by itself is often times challenging, especially at a global scale.
It requires a comprehensive dataset across the globe that would be comparable with the modelled land cover, and yet independent of the training data used to create it.
For land cover fraction mapping it is even more challenging, as fraction information needs to be available, or alternatively fine spatial resolution data from which the fractions can be calculated.
Because of these challenges, the information about the accuracy of the existing global land cover products is often limited, which in turn is problematic for the users to decide on what product suits their needs the best.
Users would also benefit from knowing the spatial variation of accuracy, as models may be more accurate at certain locations of the world and less so at others.
In this study, we made use of an independent validation set collected as part of the \gls{CGLS-LC100} project \citep{nandin-erdene_tsendbazar_copernicus_2019} that includes a total of over 20\,000 points collected across the globe, using the same legend, grid and spatial resolution as in our study.
The validation data was collected at \gls{WUR}, whereas the training data was collected at \gls{IIASA}, ensuring independence of the datasets.
We used the validation data to assess model accuracy both thematically and spatially.

@@

\begin{enumerate}
 \item Compare the performance of a variety of machine learning regression models for global-scale land cover fraction mapping
 \item Investigate approaches for tuning the models by reducing bias in the predictions with regards to zero inflation and predictions tending towards the mean
 \item Compare the importance of covariates and covariate groups for the prediction of land cover fractions
 \item Assess the accuracy of the models from a thematic and spatial point of view
\end{enumerate}

@@

The reference data (for model training and validation) used in this study was collected as part of the \ac{CGLS-LC100} project \citep{buchhorn_copernicus_2020, nandin-erdene_tsendbazar_copernicus_2019}.
The training set was generated by experts in \ac{IIASA}.
This global point dataset provides information on nine class fractions of sampled pixels of the PROBA-V 100 m product \citep{dierckx2014probav}.
The fraction estimate was acquired by high-resolution image interpretation by experts, for the year 2015.
Each of the sampled PROBA-V pixels was subdivided into a 10-by-10 m subpixel grid and the class fraction estimates were derived from the subpixel ratios.

For model validation, an independent global dataset of \ac{CGLS-LC100}, generated by \gls{WUR} \citep{tsendbazar_developing_2018}, was used.
The method of point collection is equivalent to the method used for training data collection, but performed independently by a different group of regional experts.

The data includes over 150\,000 training points and over 21\,000 validation points across the globe, describing the fractions of 12 classes in the year 2015.
The classification scheme follows the \gls{LCCS} \citep{buchhorn_copernicus_2020}.
However, due to limited number of observations for some rare classes, we merged them to get a total of seven: bare land (including snow and ice), cropland (including shifting cultivation), herbaceous vegetation (including wetland, lichen and moss), shrubs, trees, built-up and inland water.
The unknown class was discarded: points with the dominant land cover class marked as unknown were not used.
For points with a minority fraction of unknown, the remaining classes were linearly rescaled to add up to 100\%.
Thus in the end the training dataset size became 138\,164 and the validation dataset size became 20\,705. See figure \ref{fig-reference-data} for the spatial distribution of the points.

@@

\begin{figure}
 \includegraphics[width=\textwidth]{article/article-figures/maps/2020-07-06-training-and-validation}
 \caption{Sample points representing 100x100 m areas at which land cover reference data was used in the study. The colours represent the dominant land cover class at each point. (a): training dataset, collected by \gls{IIASA}, 138\,164 points used in this study. (b): validation dataset, collected by \gls{WUR}, 20\,705 points used in this study. Both datasets were collected as part of the \ac{CGLS-LC100} project \citep{buchhorn_copernicus_2020}.}
 \label{fig-reference-data}
\end{figure}

@@

% \begin{table}
% \centering
%     \begin{tabular}{llp{6cm}}
%          \toprule
%          \textbf{Category} & \textbf{Number} & \textbf{Data source} \\
%          \midrule
%          Spectral & 22 & PROBA-V 100m \ac{TOC} reflectance v1.02 \citep{probavguide2} \\
%          Harmonic & 9 & PROBA-V 100m \ac{TOC} reflectance v1.02 \citep{probavguide2} \\
%          Terrain & 4 & ASTER GDEM V003 \citep{ASTGTM003} \\
%          Climate & 21 & WorldClim 2.0 \citep{worldclim2} \\
%          Soil & 8 & SoilGrids \citep{hengl_soilgrids250m_2017} \\
%          Location & 3 & None (intrinsic) \\
%          \bottomrule
%     \end{tabular}
%     \label{tab-inputdata}
%     \caption{Data sources for the covariates input into the model.}
% \end{table}

@@

The entire archive (2014-03-11 to 2019-07-16) of the PROBA-V 100 m Level 3 Top-of-Canopy 5-day composite product \citep{dierckx2014probav,probavguide2} was used for this study.
While the reference data corresponded to the land cover at the year 2015 specifically, the whole time series of PROBA-V was chosen to obtain more robust statistics for the temporal metrics.
Land cover change is a relatively rare phenomenon and thus, according to tests performed in the making of \gls{CGLS-LC100} collection 3, is not likely to affect more than 2\% of the training points.
Nevertheless, this may result in poorer model performance compared to if reference data for each year would be available.

The time series of each of the PROBA-V spectral bands were cloud masked, first by applying the status mask provided with the product itself, and then by running a temporal cloud filter to remove the remaining outliers that were further than 2 standard deviations away from the fitted \ac{LOESS} curve over the blue reflectance band.
The resulting cloud-free time series was then used to generate the following \glspl{VI}: \gls{NDVI}, \gls{NDMI}, \gls{EVI}, \gls{OSAVI} and \gls{NIRv}. Next, the \glspl{VI} were used to calculate the descriptive statistics (median and interquartile range) over the whole time series, as well as for each phenological season separately.
Next, harmonic analysis \citep{jakubauskas2001harmonic} was run on NDVI in order to decompose the time series into sine and cosine components for two frequency orders (annual and semiannual), as well as the trend and intercept of the model.
From this, phase and amplitude for the two orders were calculated.
Lastly, the minimum and maximum of the NDVI time series was also retained as covariates.

@@

For elevation data, the ASTER GDEM v003 \citep{ASTGTM003} product (30 m) was resampled to PROBA-V 100 m grid and used as the elevation covariate.
In addition, \gls{GDAL} \citep{gdal} was used to calculate terrain parameters out of elevation: slope, aspect and \ac{TPI}.
%All terrain covariates were then resampled to the PROBA-V grid.

@@

For climate data, WorldClim 2.0 30 second product \citep{worldclim2} was used.
It includes monthly temperature, precipitation, solar radiation, wind speed and water vapour pressure data.
In addition, 19 bioclimatic parameters were calculated from the data, using the dismo package \citep{dismo}.
Some additional biophysical parameters were also calculated, namely all of the climate data during the coldest, warmest, driest and wettest months of the year at each location, as well as yearly averages of each of the original parameters.

@@

SoilGrids \citep{hengl_soilgrids250m_2017} was used for variables related to soils.
SoilGrids is based on a random forest model that predicts soil properties at various soil depths globally at a 250 m resolution.
In the creation of SoilGrids, a land cover map (based on MODIS) had been used, and so, in order to avoid circular inference, the covariates that are significantly influenced by the land cover map as detailed in \citet{hengl_soilgrids250m_2017} were excluded.
The soil taxonomy covariates were also excluded, since they are categorical derivatives from the numerical soil property data and thus do not contribute to land cover fraction prediction.

@@

In total, 313 covariates were generated in pre-processing.
However, many of them are collinear with one another, which prolongs training time for machine learning models and leads to unreliable coefficient estimation and increased error variance for linear models.
Thus, variable selection was used to remove collinear covariates before predicting the land cover fractions.
Covariates that had a Pearson's correlation coefficient $r$ \citep{pearson_notes_1895} above 0.9 were excluded in an iterative process.
After that, covariates with a Spearman's rank correlation $\rho$ \citep{spearman1904rank} above 0.9 were likewise excluded.

Most of the collinear covariates were soil covariates at different depths.
Therefore the 10 cm depth covariates were left in, and the other depths excluded, as long as $r$ was above 0.9.
Similarly, climate data was collinear between subsequent months. Thus January and July data was preferred, as these months represent different extremes of the year and are less correlated with other months' data.
In addition, all of the spectral bands of PROBA-V were also highly correlated with each other as well as to \glspl{VI}, so only \glspl{VI} were kept.

@@

For deriving land cover fractions, we compared a wide array of machine learning regression methods.
We broadly divided the methods into three types: linear models, machine learning models based on \glspl{CART}, and machine learning models not based on \glspl{CART}.
See table \ref{tab-methods} for the full list of methods that were compared in this study.
%They can be broadly divided into three categories.
%Tested linear models were \ac{FNC}, \ac{GLM}, \ac{PLS} regression, Lasso regression and \ac{MLR}.
%Non-tree machine learning models tested were \ac{MLP} \glspl{NN} and \glspl{SVM}.
%Lastly, machine learning models based on \glspl{CART} tested were Cubist and \ac{RF} regressions.
In addition, as a baseline we also compared the results with an intercept model (all fractions always predicted to be equal, namely $\frac{100\%}{7}\approx14.29\%$).

@@

\begin{table}
\centering
    \begin{tabular}{m{4cm}ll}
         \toprule
         \textbf{Category} & \textbf{Name} & \textbf{Reference} \tabularnewline
         \midrule
         Linear models &
         \begin{tabular}{@{}l@{}}
         \Glsentryfull{FNC}\tabularnewline
         \Glsentryfull{GLM}\tabularnewline
         \Glsentryfull{PLS} regression \tabularnewline
         Lasso regression \tabularnewline
         \Glsentryfull{MLR}
         \end{tabular}
         & 
         \begin{tabular}{@{}l@{}}
         \citealt{keller_fuzzy_1985} \tabularnewline
         \citealt{neter_applied_1996} \tabularnewline
         \citealt{wold_pls-regression_2001} \tabularnewline
         \citealt{tibshirani_regression_1996} \tabularnewline
         \citealt{theil_multinomial_1969}
         \end{tabular}
         \tabularnewline \midrule
         Machine learning models based on decision trees &
         \begin{tabular}{@{}l@{}}
         \Glsentryfull{RF} regression\tabularnewline
         Cubist regression
         \end{tabular}
         &
         \begin{tabular}{@{}l@{}}
         \citealt{breiman2001random} \tabularnewline
         \citealt{cubist}
         \end{tabular}
         \tabularnewline \midrule
         Other machine learning models &
         \begin{tabular}{@{}l@{}}
         \Glsentryfullpl{MLP}\tabularnewline
         \Glsentryfull{SVM} regression
         \end{tabular}
         & 
         \begin{tabular}{@{}l@{}}
         \citealt{dreyfus_artificial_1990} \tabularnewline
         \citealt{suykens_least_1999}
         \end{tabular}
         \tabularnewline \bottomrule
    \end{tabular}
    \caption{List of regression methods for land cover fraction estimation tested in this study.}
    \label{tab-methods}
\end{table}

@@

\subsection{Validation/}

@@

\end{document}
